A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
more patterns

\input{tag/counts.tex}
\input{tag/corpus.tex}
\input{tag/side-effects.tex}
\input{tag/operations.tex}
\input{tag/events.tex}
\input{tag/analysisStable.tex}
\input{tag/enrichDataset.tex}

@@

The JavaScript paper identified a number of frequently occurring eval
patterns. The patterns were detected using simple regular expressions on the
string passed to \eval. In the case of R look at the AST passed to evals.

@@

For the rest of this section we show some of the common use of \eval in R that
we have seen by looking at 20 randomly selected packages:

\subsubsection{Evaluation using Data Fields} A common use of \eval is to
  evaluate expressions that access data frame columns or list (heterogeneous
  vectors) elements by name. Core R provides two functions, \c{with} and
  \c{within} that implement this pattern. These functions construct a temporary
  environment with data frame columns or list elements as the bindings.
  Side-effects performed by the evaluated expression take place in this temporary
  environment and are ignored by the \c{with} function. The \c{within} function
  copies these modifications to a local copy of the input data and returns it.
  This pattern is used quite often for statistical modeling. The modeling
  expressions is constructed from variable names whose values are supplied by a
  data frame or list.

>>>>

\input{tag/counts.tex}
\input{tag/corpus.tex}
\input{tag/side-effects.tex}
\input{tag/operations.tex}
\input{tag/events.tex}
\input{tag/analysisStable.tex}
\input{tag/enrichDataset.tex}
\input{tag/patterns.tex}

@@

The JavaScript paper identified a number of frequently occurring \eval patterns
which were detected using simple regular expressions on the \eval strings. In
the case of R we match the patterns on the recorded AST.

Table~\ref{table:js-pattens-in-r} shows the ratio of \eval in each of these
patterns. We included all the original patterns except for \emph{JSON} as R uses
its own serialization format and \emph{empty} as there were no empty expression
passed to \eval:

@@

%\noindent
%\begin{compactitem}[$-$]
  % \item \emph{Library}: in JavaScript, the library pattern captures asynchronous
  % loading of external external function definitions. In R we have seen only
  % \PatternLibRnd such calls (\PatternLibRatio) mostly coming from source loading
  % functions (\ie \c{source} and \c{sys.source}). Most of the external function
  % definitions in R come from packages and R makes it very easy for users to
  % install packages from its centralized repository.

\noindent \emph{Library}, \emph{Typeof} and \emph{Try} are all barely used
mostly because the language provides dedicated features for it. The few calls in
Library pattern come mostly from file sourcing functions (\ie \c{source} and \c{sys.source}).
%
\noindent \emph{Read} pattern in JavaScript represented mostly an incorrect
access to an object's property or local variable both replaceable by JavaScript
core functions. In the case of R there are \PatternReadRnd \eval calls
(\PatternReadRatio). They come from calls like \c{subset(airquality, select =
  Temp)} where the variable \c{Temp} will be evaluated in the \c{airquality}
data frame.
%
\noindent \emph{Assign} category comprises of assignments to variables within
an \eval call. We have seen \PatternAssignRnd such calls (\PatternAssignRatio).
\PatternAssignArrowRatio uses the simple assignment (\c{<-}). The majority of
cases comes from the \c{magrittr}'s \c{\%<>\%} pipe operator
(\PatternAssignArrowMagrittrRatio) which forwards an object to a function on rhs
and update lhs with the result. Far more calls used the super assignment
(\c{<<-}) operator, but almost all come from the \c{plyr} package which provides
a convenient API for common data manipulation problems.
%
\noindent \emph{Call} and \emph{Other} together represents the vast majority of
eval calls in R. The \emph{Other} pattern represents all the calls where eval
was called with a value instead of with an expression. The \emph{Call} pattern
contains all the calls that did not fall in the previous categories. Without a
context it is usually difficult to further categorize the intent of these calls.
Therefore we have randomly chosen \PatternManualPackages packages and manually
looked at \PatternManualCallsites call sites from which we have observed the following categories:

% this is read pattern
% \subsubsection{Evaluation using Data Fields} A common use of \eval is to
%   evaluate expressions that access data frame columns or list (heterogeneous
%   vectors) elements by name. Core R provides two functions, \c{with} and
%   \c{within} that implement this pattern. These functions construct a temporary
%   environment with data frame columns or list elements as the bindings.
%   Side-effects performed by the evaluated expression take place in this temporary
%   environment and are ignored by the \c{with} function. The \c{within} function
%   copies these modifications to a local copy of the input data and returns it.
%   This pattern is used quite often for statistical modeling. The modeling
%   expressions is constructed from variable names whose values are supplied by a
%   data frame or list.
\subsubsection{Reconding the current call} R provides a \c{match.call} function
which can return the current call. This is heavily used by the statistical
model-fitting functions that records the current call that is then reevaluated
in the different environment. Another pattern is to match the current call,
adjust some arguments and pass it to another function. For example the
\c{write.csv} function changes the \c{sep} argument in the matched call and reevaluate itself as \c{write.table}:
%
\begin{lstlisting}
    Call <- match.call(expand.dots = TRUE)
    Call$sep <- ","; Call[[1L]] <- as.name("write.table")
    eval.parent(Call)
\end{lstlisting}
